The Evolution of Phenotypically Invariant Genetic Networks

### 2017-08-04 kalman_walk_paper.tex
removed handnotes files (moved to kalman_walk_paper) and added notes on neutral directions to kalman_walk_paper

>>>>

\subsection{Neutral directions from the Kalman decomposition}

The Kalman decomposition above says that any system $(A,B,C)$ can be decomposed into
$(P, \hat A, \hat B, \hat C)$ so that
$$\begin{aligned}
    A &= P^{-1} \hat A P  \\
    B &= P^{-1} \hat B  \\
    C &= \hat C P ,
\end{aligned}$$
and we know precisely how we can change these to preserve the transfer function:
\begin{enumerate}
    \item $P \to P + \epsilon Q$ as long as the result is still invertible,
    \item $\hat A \to A + \epsilon X$ as long as $X$ is zero in the correct places,
    \item $\hat B \to B + \epsilon Y$ as long as $Y$ is zero in the correct places,
    \item $\hat C \to C + \epsilon Y$ as long as $Z$ is zero in the correct places.
\end{enumerate}
By taking $\epsilon \to 0$, these tell us the local directions we can move a system $(A,B,C)$ in.
All statements below are up to first order in $\epsilon$,
omitting terms of order $\epsilon^2$.

First, since $(P + \epsilon Q)^{-1} = P^{-1} + \epsilon P^{-1} Q P^{-1}$,
modifying $P \to P + \epsilon Q$ changes
$$\begin{aligned}
    A 
        &\to A + \epsilon P^{-1} \hat A Q - \epsilon P^{-1} Q P^{-1} \hat A P \\
        &= A + \epsilon \left(A P^{-1} Q - P^{-1} Q A\right) , \\
    B
        &\to B - \epsilon P^{-1} Q B \\
    C
        &\to C + \epsilon C P^{-1} Q .
\end{aligned}$$

Then, $\hat A \to A + \epsilon X$  does
$$\begin{aligned}
    A \to A + \epsilon P^{-1} X P 
\end{aligned}$$
and $\hat B \to B + \epsilon Y$ does
$$\begin{aligned}
    B \to B + \epsilon P^{-1} Y
\end{aligned}$$
and $\hat C \to C + \epsilon Z$ does
$$\begin{aligned}
    C \to C + \epsilon Z P .
\end{aligned}$$

This gives us an upper bound on the number of degrees of freedom.
However, some of these may be redundant.
For instance, changing $P$ in the direction of 
a $Q$ that satisfies both $A P^{-1} Q = P^{-1} Q A$ and $C P^{-1} Q = 0$
is equivalent to changing $B$ by $Y = QB$.
